---
title: "Statistical assignment 4"
author: "[add your name here]"
date: "[add date here]"
output: github_document
---

```{r setup, include=FALSE}
# Please note these options.
# This tells R Markdown that we want to show code in the output document.
knitr::opts_chunk$set(echo = TRUE)
# Switching off messages in the output document. 
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
# Switching on caching to make things faster (don't commit cache files on Github).
knitr::opts_chunk$set(cache = TRUE)
```

In this assignment you will need to reproduce 5 ggplot graphs. I supply graphs as images; you need to write the ggplot2 code to reproduce them and knit and submit a Markdown document with the reproduced graphs (as well as your .Rmd file).

First we will need to open and recode the data. I supply the code for this; you only need to change the file paths.

    ```{r}
library(tidyverse)
    install.packages("caTools")
Data8 <- read_tsv("C:/Users/F/Documents/Politics and International relations/Data Analysis in Social Science III/UKDA-6614-tab/tab/ukhls_w8/h_indresp.tab")
Data8 <- Data8 %>%
        select(pidp, h_age_dv, h_payn_dv, h_gor_dv)
Stable <- read_tsv("C:/Users/F/Documents/Politics and International relations/Data Analysis in Social Science III/UKDA-6614-tab/tab/ukhls_wx/xwavedat.tab")
Stable <- Stable %>%
        select(pidp, sex_dv, ukborn, plbornc)
Data <- Data8 %>% left_join(Stable, "pidp")
rm(Data8, Stable)
Data <- Data %>%
        mutate(sex_dv = ifelse(sex_dv == 1, "male",
                           ifelse(sex_dv == 2, "female", NA))) %>%
        mutate(h_payn_dv = ifelse(h_payn_dv < 0, NA, h_payn_dv)) %>%
        mutate(h_gor_dv = recode(h_gor_dv,
                         `-9` = NA_character_,
                         `1` = "North East",
                         `2` = "North West",
                         `3` = "Yorkshire",
                         `4` = "East Midlands",
                         `5` = "West Midlands",
                         `6` = "East of England",
                         `7` = "London",
                         `8` = "South East",
                         `9` = "South West",
                         `10` = "Wales",
                         `11` = "Scotland",
                         `12` = "Northern Ireland")) %>%
        mutate(placeBorn = case_when(
                ukborn  == -9 ~ NA_character_,
                ukborn < 5 ~ "UK",
                plbornc == 5 ~ "Ireland",
                plbornc == 18 ~ "India",
                plbornc == 19 ~ "Pakistan",
                plbornc == 20 ~ "Bangladesh",
                plbornc == 10 ~ "Poland",
                plbornc == 27 ~ "Jamaica",
                plbornc == 24 ~ "Nigeria",
                TRUE ~ "other")
        )
    ```

Reproduce the following graphs as close as you can. For each graph, write two sentences (not more!) describing its main message.  

1. Univariate distribution (20 points).

    ```{r}
    
Data1 <- Data %>%
        filter(!is.na(h_payn_dv)) %>%
        count(h_payn_dv) 
    
Data %>%
  ggplot(aes(x = h_payn_dv, fill = "blue")) +
  geom_histogram(bins = 100) +
  xlab("net pay per month") +
  ylab("number of observations") +
  theme(legend.position = "none")

    ```

2. Line chart (20 points). The lines show the non-parametric association between age and monthly earnings for men and women.

    ```{r}
    sex_data <- Data %>%
      group_by(sex_dv) %>%
        rename(age = h_age_dv) %>%
        rename(sex = sex_dv) %>%
        rename(net_monthly_pay = h_payn_dv)
      
    ggplot(data = sex_data, 
           mapping = aes(x = h_age_dv, y = h_payn_dv, colour = sex_dv)) +
      xlab("net pay per month") +
  ylab("age (in years)") +
  
      geom_smooth(data = filter(Data))
    ```

3. Faceted bar chart (20 points).

    ```{r}
    
    sex_country <- Data %>%
        rename(Sex = sex_dv) %>%
  group_by(Sex, placeBorn) %>%
  summarise(mean_wage = mean(h_payn_dv, na.rm = TRUE)) %>% 
    filter(!is.na(Sex)) %>%
    filter(!is.na(placeBorn))

 sex_country %>%   
    ggplot(aes(x = Sex, y = mean_wage, colour = Sex)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ placeBorn, ncol = 3) +
    ylim(0,2500) +
    ylab("Mean Monthly Net Pay")
    ```

4. Heat map (20 points).

    ```{r}
    region_data <- Data %>%
  group_by(placeBorn, h_gor_dv, h_age_dv) %>%
  summarise(
    mean_age = mean(h_age_dv, na.rm = TRUE)) %>% 
    filter(!is.na(placeBorn)) %>%
      filter(!is.na(h_gor_dv))
 
 region_data %>%  
    ggplot(aes(x = h_gor_dv, y = placeBorn, fill = mean_age)) +
    geom_tile() +
    xlab("region")+
    ylab("country of origin") +
    theme(axis.text.x = element_text(angle = 90))
    
    ```

5. Population pyramid (20 points).

    ```{r}
    
Data2 <- Data %>%
        filter(!is.na(h_age_dv)) %>%
      group_by(sex_dv) %>%
        count(h_age_dv) %>%
        rename(age = h_age_dv) %>%
        rename(sex = sex_dv) %>%
        rename(population = n)
    
Data2$population <- ifelse(Data2$sex == "male",
                           -1 * Data2$population, Data2$population)    
    
ggplot(Data2, aes(x = age, y = population, fill = sex)) +
  geom_bar(subset = (Data2$sex == "female"), stat = "identity") +
  geom_bar(subset = (Data2$sex == "male"), stat = "identity") + 
  coord_flip() +
  scale_fill_brewer(palette = "Set1") + 
  theme_bw()
    
    ```